---
title: "p8105_hw1_yc4017"
author: "Elaine Yanxi Chen"
date: 2022-09-24
output: 
  github_document
---

First, we load the package needed for this assignment.

```{r load_package, message = FALSE}
library(tidyverse)
```

# Problem 1

First, we load the `penguins` dataset:

```{r import dataset}
data("penguins", package = "palmerpenguins")
```

Next, we describe the characteristics of the `penguins` dataset:

* This dataset includes the `r ncol(penguins)` variables for `r nrow(penguins)` penguins

* This dataset consists of descriptive data for `r n_distinct(penguins$species)` different species of penguins, including which island they reside on, year, sex, and information on their size such as body mass, bill length and depth, and flipper length.

* The mean flipper length for the penguins is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm

Now, we make a scatterplot and save it as a pdf:

```{r scatterplot, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species)) + geom_point()

ggsave("penguins_scatterplot.pdf")
```

# Problem 2

First, we create a data frame comprised of:

* a random sample of size 10 from a standard Normal distribution

* a logical vector indicating whether elements of the sample are greater than 0

* a character vector of length 10

* a factor vector of length 10, with 3 different factor “levels”

```{r make_df}
hw_df =
  tibble(
    var_numeric = rnorm(n = 10),
    var_logic = var_numeric > 0,
    var_char = c("Data", "science", "is", "a", "struggle", "but", "I", "really", "enjoy", "it"),
    var_factor = factor(c("easy", "medium", "hard", "medium", "medium", "easy", "hard", "easy", "hard", "hard"))
  )
```

