---
title: "p8105_hw1_yc4017"
author: "Elaine Yanxi Chen"
date: 2022-09-24
output: 
  github_document
---

First, we load the package needed for this assignment.

```{r load_package, message = FALSE}
library(tidyverse)
```

# Problem 1

First, we load the `penguins` dataset:

```{r import dataset}
data("penguins", package = "palmerpenguins")
```

Next, we describe the characteristics of the `penguins` dataset:

* This dataset includes the `r ncol(penguins)` variables for `r nrow(penguins)` penguins

* This dataset consists of descriptive data for `r n_distinct(penguins$species)` different species of penguins, including which island they reside on, year, sex, and information on their size such as body mass, bill length and depth, and flipper length.

* The mean flipper length for the penguins is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm

Now, we make a scatterplot and save it as a pdf:

```{r scatterplot, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species)) + geom_point()

ggsave("penguins_scatterplot.pdf")
```

